---
title: "Bellabeat Case Study"
author: "Emmanuel Bobmanuel"
date: "2024-07-09"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## Setting up my enviroment
Notes: setting up my R environment by loading 'tidyverse' package
```{r Loading packages, message=FALSE, warning=FALSE}
library(tidyverse)
library("ggpubr")
```


### Executive Summary

Bellabeat is a technology-driven manufacturer specializing in health-focused products for women. This case study examines the diverse range of products offered by Bellabeat and explores insights derived from analyzing data extracted from product usage. By identifying trends in customer fitness data, the company aims to unlock new growth opportunities.

### Introduction

Bellabeat, a high-tech company that manufactures health-focused smart products. This case study focuses on analyzing smart device data to gain insights into how consumers use Bella smart devices. The founders recognize growth opportunities and aim to extract valuable insights from the data. These insights will inform marketing strategies and guide the promotion of new products or ideas to enhance company growth. Specifically, we’ll explore trends in smart device usage and consider how they apply to Bellabeat customers, ultimately shaping the marketing strategy.

### Business Objective
Analyze smart device data for one of Bellabeat’s products, to gain insights into how consumers use these devices


Stakeholders

* Urška Sršen: Bellabeat’s cofounder and Chief Creative Office
* Sando Mur: Bellabeat cofounder
* Bellabeat marketing analytics team:

ABOUT DATA

This data used IS from is for Bellabeat Time Device. The Time watch connects to the Bellabeat app to provide you with insights into your daily wellness. The data is organized in a Long format. It is data collected from Thirty eligible Fitbit users who consented to the submission of personal tracker data, including minute-level output for physical activity, heart rate, and sleep monitoring

DATA STORAGE AND CREDIBILITY

The data is publicly available. It is stored in Kaggle and the dataset is made available through Mobius. The data has a Kaggle calculated score of 100% for Completeness and credibility. However there are only thirty participants and 2 months of data hence we will be focusing on the big picture to guide in our decision making and recommendations.

LICENSING, PRIVACY, SECURITY AND ACCESSIBILITY

The document is not subject to copyright. The individual who linked this work to this deed has relinquished all rights to the work worldwide under copyright law, including related and neighboring rights, to the extent permitted by law. You are free to copy, modify, distribute, and use the work, even for commercial purposes, without seeking permission.



### Data Analysis

From the datasets we have, Our analysis will encompass daily_Activities, hourly_Tracker, and sleep datasets. From this analysis, we will derive insights and formulate recommendations for management to inform our marketing approach. The data has been meticulously organized into rows and columns, covering various activities over a 2-month period. It includes information on average daily steps taken by customers, calories burned, sleep patterns, and other relevant metrics that will guide the marketing team's decision-making process.
 


```{r loading summary of dailyactivity dataset, message=FALSE, warning=FALSE}
summary(Merge_sleep_Activity <- read_csv("C:/Users/USER/Desktop/merge/Merge_sleep_Activity.csv"))
spec(hourlytracker <- read_csv("C:/Users/USER/Desktop/merge/hourlytracker.csv"))
```

DATA CLEANING

*   The data was cleaned and merged in Excel

*   The dataset arrived with 18 files with data about the participants behavior. It included activities divided into daily, hourly and by minute activities. For the analysis, we will focus on the daily and hourly activity.

*   The “sleepDay_merged” and dailyActivity_merged were merge and created as a new file called “Merge_sleep_Activity”. There were 3 duplicates which were removed. 

*   The dataset included hourly activity in 3 different files with data on Steps taken, intensities, and calories burned per hour. I combined them into 1 file named “hourlytracker”. There were no duplicates.

*   The activity date and time were in 1 column. I created a separate column for them called "Activitydate" and "time".



```{r loading code for visualization to see correlation, message=FALSE, warning=FALSE}
#We will first take at how going for a walk and the amount of steps taken daily affect the mount of calories burned vs. not being very active and having a high amount of Sedentary Minutes daily

ggarrange(
ggplot(data=Merge_sleep_Activity, aes(x=TotalSteps,y=Calories))+geom_point()+geom_smooth() +
  geom_jitter() +
  labs(title = "TotalSteps vs Calories"),
ggplot(data=Merge_sleep_Activity, aes(x=SedentaryMinutes,y=Calories))+geom_point()+geom_smooth() +
  geom_jitter() +
  labs(title = "SedentaryMinutes vs Calories")
)
```




##### Analysis

*   The data reveals a significant correlation between daily step count and calories burned. As the number of steps increases, so does the calorie expenditure. Regular walking contributes to improved physical fitness. 
*   Conversely, sedentary minutes are inversely related to calorie burn—more sedentary time leads to fewer calories expended. This underscores the importance of an active lifestyle for overall health benefits. Consistently taking more steps daily results in higher calorie burn, while prolonged inactivity negatively impacts lifestyle



```{r loading code to Average intensity and calories burned, message=FALSE, warning=FALSE}
#Here we will see how the intensity, presumably walking speed or running affect calories burned

ggplot(data=hourlytracker, aes(x= AverageIntensity, Calories))+geom_point()+geom_smooth() +
  geom_jitter() +
  labs(title = "Average Intensity vs Calories")
```




##### Analysis


*   Upon analyzing the intensity dataset, it becomes evident that as the workout or activity intensity increases, so does the number of calories burned. Consequently, to achieve greater calorie expenditure, we should elevate the intensity of our workouts. This information is crucial for determining the desired level of intensity during walks or runs



```{r loading code to check how your level of activity affects your sleep, message=FALSE, warning=FALSE}
#Here will look and how having an inactive lifestyle or not exercising and having a high amount of Sedentary Minutes could affect your sleep

ggplot(data=Merge_sleep_Activity, aes(x= SedentaryMinutes, y= TotalMinutesAsleep))+geom_point()+geom_smooth() +
  geom_jitter() +
  labs(title = "Inactivity vs Amount of Sleep")
```



##### Analysis


*   Based on our data analysis, it is clear that a sedentary lifestyle has adverse effects on sleep duration. Specifically, individuals with lower activity levels tend to experience poorer sleep quality. This insufficient sleep can also impact productivity and overall health. Therefore, it is essential to minimize sedentary minutes throughout the day to promote better sleep




```{r Want to find out what are the busiest times of day, message=FALSE, warning=FALSE}
#Here we will look at the data to see what time participant are most active during the day.
hr_tracker_new <- hourlytracker%>%
  group_by(time) %>%
  drop_na() %>%
  summarize(average_steps = mean(hourlyStepsTotal))

ggplot(data= hr_tracker_new, aes(x=time, y=average_steps)) + geom_histogram(stat = "identity", fill='blue') +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(title="Average Steps/Hr. vs. Time")
```



##### Analysis


*   The data analysis reveals that the highest step activity occurs between 8 a.m. and 7 p.m., with peak walking times concentrated from 5 to 7 p.m. This pattern likely corresponds to individuals engaging in walks or exercise during their leisure hours



### Conclusions and Recommendations.


After thorough data analysis, we have extracted valuable insights that can drive our company’s growth. The data reveals opportunities to enhance our current client experience and outlines strategies for    expanding our client base in the future. Allow me to share my recommendations for fostering growth by improving our products and services to benefit our clients

*   Develop a comprehensive program centered around combating a sedentary lifestyle. This initiative will provide valuable information on the adverse effects of inactivity, including obesity, diabetes, and cardiovascular diseases. Additionally, consider integrating the following features into your smartwatch:

    1. Sedentary Reminder: Implement a feature that alerts users when they’ve been sitting for an extended          period. The watch can vibrate or send notifications, encouraging them to move
    2. Sleep Tracking: Prioritize sleep quality by incorporating sleep-tracking capabilities. Smartwatches          can analyze sleep cycles, aiding users in managing stress and overall health.
    
*   Implement promotional campaigns or offer discounts to attract buyers. Additionally, consider strategic collaborations with third parties to enhance brand exposure. Highlight the watch’s features and their positive impact on lifestyle and overall health.

*   We should craft a compelling brand narrative that conveys the company’s origin, rooted in our unwavering passion for human health. By doing so, we aim to create an association where individuals think of our company when considering a healthy lifestyle


Thank you.
